The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
minor refinements, figures side by side

@@

\begin{figure}[hb]
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
    \caption{Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.}

\label{fig:roomba-activities}
\end{figure}

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

@@

    \caption{People's perception of the robot over time, n = 15 participants.
    This line chart shows how people's perception of the robot evolves over
    time. At 5 time points people were asked to rate the robot according to the
    following aspects: the perceived \texttt{intelligence} of the robot, the
    perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much
    \texttt{fun} it was to use the robot, how far they found the robot had an
    \texttt{impact} on their household, and their {overall} impression of it.}

    \label{fig:roomba-perception}
\end{figure}

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.7\columnwidth]{aibo-roomba-ipad_2.jpg}
%    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
%    \label{fig:aibo-roomba-ipad}
%\end{figure}

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in discussions about their technology / robot in online discussion forums.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

@@

@@

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{ranger-interactions-per-runs}
    \caption{Quantitative assessment: here, six interaction modalities have
        been selected: showing something to the robot ({\tt show}), talking to
        the robot ({\tt talk}), touching it ({\tt touch}), misbehaving (like
        shaking it, {\tt mis}), using gestures ({\tt ges}), and looking
        for help by glancing at the experimenter ({\tt look}). In each
        of the 13 experimental runs, we annotated and counted the number of
        occurrences of each of these interactions. Highlighted runs 2.3, 2.4,
        2.7 and 2.8 correspond to unexpected robot behaviors: we observe peaks
        of interactions every time the robot's behavior changes from expected to
        unexpected or from unexpected to expected. (n = 26 children)}

    \label{fig:interactions-per-run}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.5\columnwidth]{ranger-interactions-vs-qualitative-score}
    \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. This 
    suggests that children who interacted more with the robot, anthropomorphized it 
    less, and \textit{vice versa}. (n = 13 groups)}
    \label{fig:qualitative-score}
\end{figure}

@@

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it appears controversial, at 
the same time, that humans tend to \hl{\textit{``see human in the non-
human''}}. 
Anthropomorphism as a social phenomenon arises in an interaction 
\citep{persson_anthropomorphism_2000}, and thus includes at least two interaction 
partners (the human and the real or imagined non-human agent). Therefore, one can 
take two main perspectives on explaining it \citep{lee_human_2005}: an artifact-
centered, and a human-centered. We briefly outline both views in the following. 
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same kind of phenomenon.

In this section we will try to create a common ground in the understanding of 
anthropomorphism, with a focus on it related to human-robot interaction. 
First of all, we would like to propose a clarification of terms.
We call an imitation of human form (or in a broader sense also animal and natural) 
form \textit{``anthropomorphic design''} or \textit{``anthropomorphic form''} -- 
but \textit{not anthropomorphism}. Contrary, we call the social phenomenon of 
people's tendency to perceive human-likeness in non human-like 
agents \textit{anthropomorphism}.

In the following, we first present an artifact-centered perspective on 
anthropomorphism, and outline how the human-like 
design approach (\textit{anthropomorphic form}) is used in robotics to encourage 
people to interact socially with robots. Then, we apply a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from psychology and cognitive science to explain people's tendency to 
perceive human-like characteristics in non-human agents. In the end, we synthesize 
the presented backgrounds and suggest an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts / the design of the non-human agent.
The direction of causality in this view is that the non-human artifact (or certain 
features of it) encourages the human to anthropomorphize it.
This perspective assumes that humans directly respond to life-like and social cues 
that the non-human artifact or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it. In fact, from early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
and to respond to social cues, such as facial expressions, social gestures or turn-
taking. When first looking at an unknown object, the visual scope of its design, 
its outer appearance is likely to have an important impact on how the object is 
overall perceived (how it appears to the human) \citep{schmitz_concepts_2011}.

@@

However, the anthropomorphic form also 
increases people's expectations of the agent, and consequently, a very human-like 
appearing robot is expected to act like a human. But today's artificial systems are 
not able (yet) to mimic human movements, or human social cues in an acceptable way. 
The case is that the artificial agent pretends to be human but fails, and this is 
more than just disappointing for the 
human -- it appears uncanny and evokes reluctance, the so-called 
\textit{"uncanny valley"}. The idea of naming this shift and drop from a positive to a negative reaction,
follows Freud's description of the so-called \textit{uncanny} (a translation from the German word 
\textit{``unheimlich''}). Controversially, something uncanny \textit{"derives its terror not from 
something externally alien or unknown but -- on the contrary -- from something 
strangely familiar which defeats our efforts to separate ourselves from it"} (cited 
after \cite{hegel_understanding_2008}).
In other words, humans are very sensitive to the limits of what is familiar to them 
in a sense of being human. At the same time, humans are naturally very skilled when 
it comes to identifying what is acceptable as a human and what is a \hl{NACHAHMUNG} 
of it. Even subtle differences can be crucial -- humans are inherently well trained 
in identifying if there is \hl{``something strange''} in the familiar.
Summing it up, a not yet very well defined thin line separates acceptable human-like forms in 
robots from creepy artificial \textit{almost} human-like appearing robots.

>>>>

\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\floatstyle{plaintop}

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\columnwidth]{roomba-activities.png}
%    \caption{Activities with the robot over time, n = 9 households. This bar
%        chart shows how many of the 9 households carried out different kinds of
%        anthropomorphic / social activities with and related to the robot at 3
%        time points during 6 months. Activities include \texttt{watching} the
%        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
%        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
%        \texttt{talking} to the robot, \texttt{playing} with the robot.
%        Participants were asked to report on these activities at 3 time points
%        after the robot was deployed in their home: after 2 weeks, after 2 months,
%        and after 6 months.} 
%
%\label{fig:roomba-activities}
%\end{figure}

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Roomba activities}
                \label{fig:roomba-activities}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{Roomba perception}
                \label{fig:roomba-perception}
        \end{subfigure}
        \caption{\textbf{The Roomba study: Long-term interaction with a robot.\\}
    \textbf{a)} Activities with the robot over time, n = 9 households. This bar
        chart shows how many of the 9 households carried out different kinds of
        anthropomorphic / social activities with and related to the robot at 3
        time points during 6 months. Activities include \texttt{watching} the
        robot for fun, giving a \texttt{demo} of the robot to guests / visitors,
        \texttt{personalizing} the robot, giving a \texttt{name} to the robot,
        \texttt{talking} to the robot, \texttt{playing} with the robot.
        Participants were asked to report on these activities at 3 time points
        after the robot was deployed in their home: after 2 weeks, after 2 months,
        and after 6 months.\\
	\textbf{b)} People's perception of the robot over time, n = 15 participants.
    		This line chart shows how people's perception of the robot evolves over
    		time. At 5 time points people were asked to rate the robot according to the
    		following aspects: the perceived \texttt{intelligence} of the robot, the
    		perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much
    		\texttt{fun} it was to use the robot, how far they found the robot had an
    		\texttt{impact} on their household, and their {overall} impression of it.}
    		\label{fig:roomba-study}
\end{figure}

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\columnwidth]{roomba-perception.pdf}

@@

%    \caption{People's perception of the robot over time, n = 15 participants.
%    This line chart shows how people's perception of the robot evolves over
%    time. At 5 time points people were asked to rate the robot according to the
%    following aspects: the perceived \texttt{intelligence} of the robot, the
%    perceived \texttt{usefulness}, the perceived \texttt{ease of use}, how much
%    \texttt{fun} it was to use the robot, how far they found the robot had an
%    \texttt{impact} on their household, and their {overall} impression of it.}
%
%    \label{fig:roomba-perception}
%\end{figure}

@@

@@

@@

\begin{figure}[b]
    \centering
    \includegraphics[width=0.5\columnwidth]{aibo-roomba-ipad_2.jpg}
    \caption{The Sony AIBO robotic dog (left), the iRobot Roomba vacuum-cleaning robot (middle), and the Apple Inc. iPad tablet computer (right) were compared in terms of how far users of these devices use anthropomorphic expressions in online discussion forums about their technology / robot.}
    \label{fig:aibo-roomba-ipad}
\end{figure}

@@

\begin{figure}
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{ranger-interactions-per-runs}
                \caption{Interactions per run}
                \label{fig:interactions-per-run}
        \end{subfigure}%
        ~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %(or a blank line to force the subfigure onto a new line)
        \begin{subfigure}[b]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
                \caption{Qualitative score by interactions}
                \label{fig:qualitative-score}
        \end{subfigure}
        \caption{\textbf{The Ranger domino study: Child-robot interaction.}\\
    \textbf{a)} Quantitative assessment: here, six interaction modalities have
        been selected: showing something to the robot ({\tt show}), talking to
        the robot ({\tt talk}), touching it ({\tt touch}), misbehaving (like
        shaking it, {\tt mis}), using gestures ({\tt ges}), and looking
        for help by glancing at the experimenter ({\tt look}). In each
        of the 13 experimental runs, we annotated and counted the number of
        occurrences of each of these interactions. Highlighted runs 2.3, 2.4,
        2.7 and 2.8 correspond to unexpected robot behaviors: we observe peaks
        of interactions every time the robot's behavior changes from expected to
        unexpected or from unexpected to expected. (n = 26 children)\\
	\textbf{b)} We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. This 
    suggests that children who interacted more with the robot, anthropomorphized it 
	less, and \textit{vice versa}. (n = 13 groups)}
    		\label{fig:Ranger-domino}
\end{figure}

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\columnwidth]{ranger-interactions-per-runs}
%    \caption{Quantitative assessment: here, six interaction modalities have
%        been selected: showing something to the robot ({\tt show}), talking to
%        the robot ({\tt talk}), touching it ({\tt touch}), misbehaving (like
%        shaking it, {\tt mis}), using gestures ({\tt ges}), and looking
%        for help by glancing at the experimenter ({\tt look}). In each
%        of the 13 experimental runs, we annotated and counted the number of
%        occurrences of each of these interactions. Highlighted runs 2.3, 2.4,
%        2.7 and 2.8 correspond to unexpected robot behaviors: we observe peaks
%        of interactions every time the robot's behavior changes from expected to
%        unexpected or from unexpected to expected. (n = 26 children)}
%
%    \label{fig:interactions-per-run}
%\end{figure}

@@

%\begin{figure}
%    \centering
%    \includegraphics[width=0.5\columnwidth]{ranger-interactions-vs-qualitative-score}
%    \caption{We found a significant negative correlation ($p=0.048$) between the
%    number of interactions and the qualitative anthropomorphic projections. This 
%    suggests that children who interacted more with the robot, anthropomorphized it 
%    less, and \textit{vice versa}. (n = 13 groups)}
%    \label{fig:qualitative-score}
%\end{figure}

@@

Anthropomorphism is an interesting phenomenon because it tells us something about 
the relationship between the human who anthropomorphizes, and the non-human agent 
which is anthropomorphized. Though commonly observed, it appears controversial, at 
the same time, that humans tend to \textit{``see human in the non-
human''} \citep{epley_seeing_2007}. 
Anthropomorphism as a social phenomenon arises in an interaction 
\citep{persson_anthropomorphism_2000}, and thus includes at least two interaction 
partners: the human and the real or imagined non-human agent. Therefore, one can 
take two main perspectives on explaining anthropomorphism \citep{lee_human_2005}: a human-centered and an artifact-centered. We briefly outline both views in the following. 
It has to be kept in mind that both perspectives are inseparable; they apply at the 
same time, and we would like to present them here just as two different viewpoints 
on the same phenomenon.

By doing so, this section tries to create a common ground in the understanding of 
anthropomorphism, with a focus on the design of robots and human-robot interaction. 
First of all, we would like to propose a clarification of terms.
In short, we propose to use two different terms, in order to differentiate between the human-like design of a non-human agent, and the perception of human-likeness in a non-human agent.
We use the terms \textit{``anthropomorphic design''} and \textit{``anthropomorphic 
form''} interchangeably to refer to an imitation of human form (or in a broader sense 
also animal and natural) in the design of objects and artificial agents. We propose to 
\textit{not} refer to this as \textit{anthropomorphism}.
Because we use \textit{anthropomorphism} to refer to the social phenomenon of 
people's tendency to perceive human-likeness in non human-like agents. This is an important differentiation.

In the following, we first present an artifact-centered perspective on anthropomorphism, and outline how 
the human-like design approach (\textit{anthropomorphic form}) is used in robotics to encourage 
people to interact in a social way with robots. Then, we apply a human-centered 
perspective and shed light on \textit{anthropomorphism as a social phenomenon}. We present 
theories from psychology and cognitive science to explain people's tendency to 
perceive human-like characteristics in non-human agents. In the end, we synthesize 
the presented backgrounds and suggest an extended understanding of anthropomorphism 
as a multi-factor phenomenon.

@@

An artifact-centered perspective explains anthropomorphism from the ascertainable 
parts / the design of the non-human agent.
The direction of causality in this view is that the non-human artifact (or a certain 
feature of it) encourages the human to anthropomorphize the artifact / agent.
This perspective assumes that humans directly respond to life-like and social cues 
that the non-human agent or system emits, without thoughtful mental processing, 
by simply applying stereotypes and heuristics to it. In fact, from early childhood 
on, humans are inherently well-trained to perceive life \citep{epley_seeing_2007}, 
and to respond to social cues, such as facial expressions, social gestures or turn-
taking. When first looking at an unknown object, the visual scope of its design, 
its outer appearance is likely to have an important impact on how the object is 
overall perceived (how it appears to the human) \citep{schmitz_concepts_2011}.

@@

However, the anthropomorphic form also 
increases people's expectations of the agent, and consequently, a very human-like 
appearing robot is expected to act like a human. But today's artificial systems are 
not able (yet) to mimic human movements, or human social cues in an acceptable way. 
The case is that the artificial agent pretends to be human but fails, and this is 
more than just disappointing for the 
human -- it appears uncanny and evokes reluctance, the so-called 
\textit{"uncanny valley"}. The idea of naming this shift and drop from a positive to a negative reaction,
follows Freud's description of the so-called \textit{uncanny} (a translation from the German word 
\textit{``unheimlich''}). Controversially, something uncanny \textit{"derives its terror not from 
something externally alien or unknown but -- on the contrary -- from something 
strangely familiar which defeats our efforts to separate ourselves from it"} (cited 
after \cite{hegel_understanding_2008}).
In other words, humans are very sensitive to the limits of what is familiar to them 
in a sense of being human. At the same time, humans are naturally very skilled when 
it comes to identifying what is acceptable as a human and what is an imitation 
of it. Even subtle differences can be crucial -- humans are inherently well trained 
in identifying if there is \hl{``something strange''} in the familiar.
Summing it up, a not yet very well defined thin line separates acceptable human-like forms in 
robots from creepy artificial \textit{almost} human-like appearing robots.
